---
title: "Identifying Discrepencies and Predictors of Early Academic Achievement Across the United States"
subtitle: "STA 113FS - Fall 2023 - Project 2"
author: "JanIce"
format: 
  revealjs:
    theme: simple
editor: visual
execute:
  echo: false
---

```{r}
#| label: loading-packages
#| message: false 

library(tidyverse)
library(tidymodels)
library(maps)
library(leaps)
library(dplyr)
library(scales)
```

```{r}
#| label: loading-datasets
#| message: false

#loading datasets childcare_costs and counties from tidytuesday

#loading dataset childcare_costs
childcare_costs <- read_csv("data/childcare_costs.csv")

#loading dataset counties
counties <- read_csv("data/counties.csv")

#loading dataset expenditures
expenditures <- read_csv("data/expenditures.csv")

#loading dataset scores
scores <- read_csv("data/scores.csv")

#loading us map 
us_states <- map_data("state")
```

## Contextualizing Our Project

-   Four courses in KISS connected by two learning goals:

    -   Root cause analysis: elucidating roots of issues to develop meaningful solutions to sociopolitical issues

    -   Tools for interpersonal connection and collective advocacy

## Project Objectives

1.  Illustrate relative average 4th-grade reading level in each state

2.  Provide side-by-side comparison of funding each state allocates to students

3.  Construct best-fit linear regression model with community-level factors associated with average 4th-grade reading level across a county

4.  Determine distribution of occupational fields in states with the best 4th-grade reading levels

## Choropleth Map: 2022 Scores

```{r}
#| label: us-map-scores
#| fig.alt: "This visualization is a choropleth map of the United States, a plot that indicates the strength of a certain variable across different states by increasing or decreasing their shading (excluding Hawaii and Alaska). This plot has the title 'Achievement Rates Based on State' and is mapping 2016 4th-grade reading achievement rates from 2022 to each state. From this visualization, we see that there does not appear to be a clear pattern in higher achievement rates--Massachusetts, New Jersey, and Florida seem to be the highest achieving states at around 225 while New Mexico and West Virginia appear to be the lowest achieving ones (at around 200). The states of Wyoming, Utah, and Colorado are also high-achieving (with scores of about 225, similarly to Massachusetts) among many other average- or low-achieving states."

scores <- scores|> 
  rename(reading_level = MN)

scores_new <- scores |> 
  filter(reading_level != is.na(reading_level)) |>
  mutate(region = str_to_lower(Jurisdiction), reading_level = as.numeric(reading_level))

us_states_scores <- us_states |>
  select(-subregion) |>
  inner_join(
    scores_new, 
    join_by(region == region), 
    relationship = "many-to-many"
    )

ggplot(us_states_scores, aes(x = long, y = lat, group = group, fill = reading_level)) +
  geom_polygon(color = "white") + 
  theme_void() + 
  scale_fill_gradient(low = "gray90", high = "darkblue") +
  theme(
    legend.text = element_text(size=7),
    legend.title = element_text(size=10),
    legend.position = c(.2, .1),
    legend.direction = "horizontal"
        ) +
  coord_quickmap() + 
  labs(
    title = "Achievement Rates Based on State", 
    fill = "Average 4th Grade \nReading Score"
    )
```

## Choropleth Map: 2016 Expenditures

```{r}
#| label: us-map-dataset

expenditures_PK12ed <- expenditures |> 
  mutate(region = str_to_lower(state)) |>
  filter(variable == "PK12ed")

us_states_expenditures <- us_states |>
  select(-subregion) |>
  inner_join(
    expenditures_PK12ed, 
    join_by(region == region), 
    relationship = "many-to-many"
    ) |> 
  filter(region != "district of columbia") |> 
  filter(year == 2016)
```

```{r}
#| label: us-map-expenditures 
#| fig.alt: "This visualization is a choropleth map of the United States, a plot that indicates the strength of a certain variable across different states by increasing or decreasing their shading (excluding Hawaii and Alaska). This plot has the title 'Expenditures Based on State' and is mapping 2016 state expenditures on education per child, adjusted for inflation, into each state. From this visualization, we see that states in the Northeast, namely New York, New Jersey, and Vermont, appear to provide the most funding to their students (at around $13,000). We can also see that states in the South and West, specifically New Mexico, Idaho, and Utah, have the lowest expenditure per child (at around $2,500). Most of the Midwest appears to have average per child expenditures, but Wyoming stands out as having significantly greater amounts allotted towards this."

ggplot(
  us_states_expenditures, 
  aes(
    x = long, 
    y = lat, 
    group = group, 
    fill = inf_adj_perchild
    )
  ) +
  geom_polygon(color = "white") +
  theme_void() +
  theme(
    legend.text = element_text(size=7),
    legend.title = element_text(size=10),
    legend.position = c(.2, .1),
    legend.direction = "horizontal"
    ) + 
  scale_fill_gradient(low = "gray90", high = "darkblue") +
  coord_quickmap() + 
  labs(
    title = "Expenditures Based on State", 
    fill = "Expenditure \nper Child \n(in thousands)"
    )
```

## Linear Model Literature Review:

-   Behavioral skills influenced by parental monitoring and family stability (EARLY SKILLS AND PREDICTORS OF ACADEMIC SUCCESS 2016)

    -   Crucial behavioral skills predict reading level

-   Low-SES families less likely to have experiences developing skills such as phonological awareness, vocabulary, and oral language (American Psychological Association 2017)

## Linear Model Literature Review (cont'd):

-   Longitudinal study of 370+ working-class families finds jobs allow for more autonomy, and supportive supervisors were more engaged with their children (Perry-Jenkins 2023)

-   Achievement gap amongst students of varying races particularly between White, Black, and Hispanic children (Racial and Ethnic Achievement Gaps n.d.)

    -   Likely mediated by differences in the socioeconomic statuses of families

## Linear Model Literature Review (cont'd):

-   High-quality childcare has been shown to lead to better performance in STEM through high school (American Psychological Association 2023) 

```{r}
#| label: merging-data 

#Merging childcare_costs and counties 
merged_counties_data <- left_join(childcare_costs, counties, join_by("county_fips_code"))

#Preparing scores to merge
renamed_scores <- scores |> 
  rename("state_name" = "Jurisdiction")

#Merging all data
merged_data <- left_join(merged_counties_data, renamed_scores, join_by("state_name"))
```

```{r}
#| label: preparing-data-for-vis-3
#| echo: false

#Preparing data
best_states_data <- merged_data |> 
  arrange(desc(reading_level)) |>
  mutate(achievement_level = if_else(reading_level>=223, "High", "Low")) |>
  filter(achievement_level == "High") |> 
  filter(!is.na(state_name))

best_states_data <- best_states_data |>
  relocate(state_name) |> 
  pivot_longer(
    cols = c("emp_m", "emp_service", "emp_sales", "emp_n", "emp_p"), 
    names_to = "sector_type", 
    values_to = "sector_percentage"
  ) |>
  mutate(
    sector_type = case_when(
      sector_type == "emp_m" ~ "Management and Science", 
      sector_type == "emp_n" ~ "Natural Resources", 
      sector_type == "emp_p" ~ "Production", 
      sector_type == "emp_service" ~ "Public Service",
      sector_type == "emp_sales" ~ "Sales and Office"
    )
  )

#Merging childcare_costs and counties    
merged_counties_data <- left_join(
  childcare_costs, 
  counties, 
  join_by("county_fips_code")
  )

#Preparing scores to merge
renamed_scores <- scores |> 
  rename("state_name" = "Jurisdiction")

#Merging all data
merged_data <- left_join(
  merged_counties_data, 
  renamed_scores, 
  join_by("state_name")
  )
```

```{r}
#| label: forward-selection-model-variables

#Family poverty level 
merged_data <- merged_data |>
  mutate(family_poverty = pr_f)

#Minority racial status 
merged_data <- merged_data |> 
  mutate(
    minority_racial_status = 
      one_race_b +
      one_race_i +
      one_race_a +
      one_race_h +
      one_race_other + 
      two_races
  )

#Percentage of single-mother households 
merged_data <- merged_data |> 
  mutate(percentage_single_mother = h_under6_single_m / households)
  
#Ratio of one-parent-working to two-parent-working households 
merged_data <- merged_data |> 
  mutate(
    one_parent_households = h_under6_f_work + h_under6_f_work, 
    ratio_of_one_to_both = one_parent_households / h_under6_both_work
  )

#Cost of center-based child care
merged_data <- merged_data |> 
  mutate(
    average_center_child_care = 
    (mc_infant + mc_toddler + mc_preschool) / 3
  )

#Cost of family-based child care 
merged_data <- merged_data |> 
  mutate(
    average_family_child_care = 
    (mfcc_infant + mfcc_toddler + mfcc_preschool) / 3
  )
```

## Final Regression Model (#61)

```{r}
#| label: forward-selection-model 

#filtering by year 
merged_data_2018 <- merged_data |> 
  filter(study_year == "2018")

#Creating regression model with all possible variables 
merged_data_2018 <- merged_data_2018 |> 
  filter(
    !is.na(reading_level),
    !is.na(family_poverty),
    !is.na(percentage_single_mother),
    !is.na(ratio_of_one_to_both),
    !is.na(average_center_child_care),
    !is.na(average_family_child_care),
    !is.na(one_parent_households),
    ratio_of_one_to_both != 'Inf') |>
  mutate(reading_level = as.numeric(reading_level)
         ) 

maximum_regression_model <- linear_reg() |>
  fit(
    reading_level ~ 
      family_poverty + 
      percentage_single_mother + 
      average_center_child_care + 
      average_family_child_care + 
      one_parent_households + 
      ratio_of_one_to_both, 
    data = merged_data_2018
    )

regression_model_1 <- linear_reg() |> 
  fit(reading_level ~ family_poverty, data = merged_data_2018)
```

```{r}
#| label: preparing-data-for-model
#| echo: false

#Preparing data
best_states_data <- merged_data_2018 |> 
  arrange(desc(reading_level)) |>
  mutate(achievement_level = if_else(reading_level>=223, "High", "Low")) |>
  filter(achievement_level == "High")

best_states_data <- best_states_data |>
  relocate(state_name) |> 
  pivot_longer(
    cols = c("emp_m", "emp_service", "emp_sales", "emp_n", "emp_p"), 
    names_to = "sector_type", 
    values_to = "sector_percentage"
  ) |>
  mutate(sector_type = case_when(
    sector_type == "emp_m" ~ "Management and Science", 
    sector_type == "emp_n" ~ "Natural Resources", 
    sector_type == "emp_p" ~ "Production", 
    sector_type == "emp_service" ~ "Public Service",
    sector_type == "emp_sales" ~ "Sales and Office"
    ), 
    state_name = case_when(
      state_name == "Massachusetts" ~ "MA", 
      state_name == "New Hampshire" ~ "NH",
      state_name == "New Jersey" ~ "NJ",
      state_name == "Wyoming" ~ "WY"
    )
  )

#Grouping and summarizing data 
summarized_data <- best_states_data |>
  group_by(state_name, sector_type) |>
  summarize(mean_percentage = mean(sector_percentage)) 
```

```{r}
#| label: creating-all-regression-models

regression_model_1 <- lm(
  reading_level ~ 
    family_poverty, 
  data = merged_data_2018
  )

regression_model_2 <- lm(
  reading_level ~ 
    percentage_single_mother, 
  data = merged_data_2018
  )

regression_model_3 <- lm(
  reading_level ~ 
    average_center_child_care, 
  data = merged_data_2018
  )

regression_model_4 <- lm(
  reading_level ~ 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_5 <- lm(
  reading_level ~ 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_6 <- lm(
  reading_level ~ 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_7 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother, 
  data = merged_data_2018
  )

regression_model_8 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care, 
  data = merged_data_2018
  )

regression_model_9 <- lm(
  reading_level ~ 
    family_poverty + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_10 <- lm(
  reading_level ~ 
    family_poverty + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_11 <- lm(
  reading_level ~ 
    family_poverty + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_12 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care, 
  data = merged_data_2018
  )

regression_model_13 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_14 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_15 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_16 <- lm(
  reading_level ~ 
    average_center_child_care + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_17 <- lm(
  reading_level ~ 
    average_center_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_18 <- lm(
  reading_level ~ 
    average_center_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_19 <- lm(
  reading_level ~ 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_20 <- lm(
  reading_level ~ 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_21 <- lm(
  reading_level ~ 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_22 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care, 
  data = merged_data_2018
  )

regression_model_23 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_24 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_25 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_26 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_27 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_28 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_29 <- lm(
  reading_level ~ 
    family_poverty +
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_30 <- lm(
  reading_level ~ 
    family_poverty + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_31 <- lm(
  reading_level ~ 
    family_poverty + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_32 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_33 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_34 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_35 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_36 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_37 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_38 <- lm(
  reading_level ~ 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_39 <- lm(
  reading_level ~ 
    average_center_child_care + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_40 <- lm(
  reading_level ~ 
    average_center_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_41 <- lm(
  reading_level ~ 
    average_family_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_42 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care, 
  data = merged_data_2018
  )

regression_model_43 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_44 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_45 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_46 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_47 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_48 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_49 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_50 <- lm(
  reading_level ~ 
    family_poverty + 
    average_center_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_51 <- lm(
  reading_level ~ 
    family_poverty + 
    average_family_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_52 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_53 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_54 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_center_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_55 <- lm(
  reading_level ~ 
    percentage_single_mother + 
    average_family_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_56 <- lm(
  reading_level ~ 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_57 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_58 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_59 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )

regression_model_60 <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_family_child_care + 
    ratio_of_one_to_both + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_61 <- lm(
  reading_level ~ 
    family_poverty + 
    ratio_of_one_to_both + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

regression_model_62 <- lm(
  reading_level ~ 
    ratio_of_one_to_both + 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households, 
  data = merged_data_2018
  )

maximum_regression_model <- lm(
  reading_level ~ 
    family_poverty + 
    percentage_single_mother + 
    average_center_child_care + 
    average_family_child_care + 
    one_parent_households + 
    ratio_of_one_to_both, 
  data = merged_data_2018
  )
```

```{r}
#| label: creating-dataframe-analyzing-model
#| echo: false

r_squared <- c(summary(regression_model_1)$adj.r.squared, 
          summary(regression_model_2)$adj.r.squared, 
          summary(regression_model_3)$adj.r.squared, 
          summary(regression_model_4)$adj.r.squared, 
          summary(regression_model_5)$adj.r.squared, 
          summary(regression_model_6)$adj.r.squared, 
          summary(regression_model_7)$adj.r.squared, 
          summary(regression_model_8)$adj.r.squared, 
          summary(regression_model_9)$adj.r.squared, 
          summary(regression_model_10)$adj.r.squared, 
          summary(regression_model_11)$adj.r.squared, 
          summary(regression_model_12)$adj.r.squared, 
          summary(regression_model_13)$adj.r.squared, 
          summary(regression_model_14)$adj.r.squared, 
          summary(regression_model_15)$adj.r.squared, 
          summary(regression_model_16)$adj.r.squared, 
          summary(regression_model_17)$adj.r.squared, 
          summary(regression_model_18)$adj.r.squared, 
          summary(regression_model_19)$adj.r.squared, 
          summary(regression_model_20)$adj.r.squared, 
          summary(regression_model_21)$adj.r.squared, 
          summary(regression_model_22)$adj.r.squared, 
          summary(regression_model_23)$adj.r.squared, 
          summary(regression_model_24)$adj.r.squared, 
          summary(regression_model_25)$adj.r.squared, 
          summary(regression_model_26)$adj.r.squared, 
          summary(regression_model_27)$adj.r.squared, 
          summary(regression_model_28)$adj.r.squared, 
          summary(regression_model_29)$adj.r.squared, 
          summary(regression_model_30)$adj.r.squared, 
          summary(regression_model_31)$adj.r.squared, 
          summary(regression_model_32)$adj.r.squared,
          summary(regression_model_33)$adj.r.squared, 
          summary(regression_model_34)$adj.r.squared, 
          summary(regression_model_35)$adj.r.squared, 
          summary(regression_model_36)$adj.r.squared, 
          summary(regression_model_37)$adj.r.squared, 
          summary(regression_model_38)$adj.r.squared, 
          summary(regression_model_39)$adj.r.squared, 
          summary(regression_model_40)$adj.r.squared, 
          summary(regression_model_41)$adj.r.squared, 
          summary(regression_model_42)$adj.r.squared, 
          summary(regression_model_43)$adj.r.squared, 
          summary(regression_model_44)$adj.r.squared, 
          summary(regression_model_45)$adj.r.squared, 
          summary(regression_model_46)$adj.r.squared, 
          summary(regression_model_47)$adj.r.squared, 
          summary(regression_model_48)$adj.r.squared,
          summary(regression_model_49)$adj.r.squared, 
          summary(regression_model_50)$adj.r.squared, 
          summary(regression_model_51)$adj.r.squared, 
          summary(regression_model_52)$adj.r.squared, 
          summary(regression_model_53)$adj.r.squared, 
          summary(regression_model_54)$adj.r.squared, 
          summary(regression_model_55)$adj.r.squared, 
          summary(regression_model_56)$adj.r.squared, 
          summary(regression_model_57)$adj.r.squared, 
          summary(regression_model_58)$adj.r.squared, 
          summary(regression_model_59)$adj.r.squared, 
          summary(regression_model_60)$adj.r.squared, 
          summary(regression_model_61)$adj.r.squared, 
          summary(regression_model_62)$adj.r.squared, 
          summary(maximum_regression_model)$adj.r.squared)

model_name <- c("regression_model_1",
                  "regression_model_2",
                  "regression_model_3",
                  "regression_model_4",
                  "regression_model_5",
                  "regression_model_6",
                  "regression_model_7",
                  "regression_model_8",
                  "regression_model_9",
                  "regression_model_10",
                  "regression_model_11",
                  "regression_model_12",
                  "regression_model_13",
                  "regression_model_14",
                  "regression_model_15",
                  "regression_model_16",
                  "regression_model_17",
                  "regression_model_18",
                  "regression_model_19",
                  "regression_model_20",
                  "regression_model_21",
                  "regression_model_22",
                  "regression_model_23",
                  "regression_model_24",
                  "regression_model_25",
                  "regression_model_26",
                  "regression_model_27",
                  "regression_model_28",
                  "regression_model_29",
                  "regression_model_30",
                  "regression_model_31",
                  "regression_model_32",
                  "regression_model_33",
                  "regression_model_34",
                  "regression_model_35",
                  "regression_model_36",
                  "regression_model_37",
                  "regression_model_38",
                  "regression_model_39",
                  "regression_model_40",
                  "regression_model_41",
                  "regression_model_42",
                  "regression_model_43",
                  "regression_model_44",
                  "regression_model_45",
                  "regression_model_46",
                  "regression_model_47",
                  "regression_model_48",
                  "regression_model_49",
                  "regression_model_50",
                  "regression_model_51",
                  "regression_model_52",
                  "regression_model_53",
                  "regression_model_54",
                  "regression_model_55",
                  "regression_model_56",
                  "regression_model_57",
                  "regression_model_58",
                  "regression_model_59",
                  "regression_model_60",
                  "regression_model_61",
                  "regression_model_62",
                  "regression_model_63")

final <- data.frame(model_name, r_squared)

final_model <- final |>
  filter(r_squared == max(r_squared))
```

```{r}
#| label: model-tidy-output

tidy(regression_model_61) |> 
  knitr::kable(digits = 3)
```

## Job Distributions in Best-Practice States

```{r}
#| label: preparing-for-vis-4

#Preparing data
best_states_data <- merged_data |> 
  arrange(desc(reading_level)) |>
  mutate(achievement_level = if_else(reading_level>=223, "High", "Low")) |>
  filter(achievement_level == "High") |> 
  filter(!is.na(state_name))

best_states_data <- best_states_data |>
  relocate(state_name) |> 
  pivot_longer(
    cols = c("emp_m", "emp_service", "emp_sales", "emp_n", "emp_p"), 
    names_to = "sector_type", 
    values_to = "sector_percentage"
  ) |>
  mutate(sector_type = case_when(
    sector_type == "emp_m" ~ "Management and Science", 
    sector_type == "emp_n" ~ "Natural Resources", 
    sector_type == "emp_p" ~ "Production", 
    sector_type == "emp_service" ~ "Public Service",
    sector_type == "emp_sales" ~ "Sales and Office"
    ), 
    state_name = case_when(
      state_name == "Massachusetts" ~ "MA", 
      state_name == "New Hampshire" ~ "NH",
      state_name == "New Jersey" ~ "NJ",
      state_name == "Wyoming" ~ "WY"
    )
  )

#Grouping and summarizing data 
summarized_data <- best_states_data |>
  group_by(state_name, sector_type) |>
  summarize(mean_percentage = mean(sector_percentage))


#Trying somethign new
summarized_data_1 <- best_states_data |>
  group_by(county_fips_code, study_year) |>
  summarize(sum_percentage = sum(sector_percentage))

summarized_data_2 <- best_states_data |>
  group_by(state_name, sector_type, study_year) |>
  summarize(mean_percentage = mean(sector_percentage))
  
#Merging data 
merged_intermediary_data <- left_join(best_states_data, summarized_data_1, join_by(county_fips_code, study_year))

merged_final_data <- left_join(merged_intermediary_data, summarized_data_2, join_by(state_name, sector_type, study_year))
```

```{r}
#| label: visualizing-states
#| fig-asp: 0.6
#| fig.alt: "The visualization is a density plot showing the overlaid breakdown of different sectors for six states of interest. Its title is “Percentage of Population Employed in Sector by State.” Its legend is called “Sector.” There are four plots shown stacked beside and on top of each other, each for a different state. Those states are Massachusetts, New Hampshire, New Jersey, Florida, and Colorado and Wyoming. The x-axis is the year when a datapoint was obtained (from 2000 to 2018), and the y-axis is the percentage of the population employed by each respective sector (from 0 to 40%). Each plot except for that of Wyoming illustrates the following: management and science having most employed in management and science (around 40%), the second-most (under 30%) employed by sales and office, the third-most (around 20%) employed in public service, and the least (around 10% each) employed by the two industries of natural resources and production. Wyoming differs from the other more clearly different plots because the majority of WY (around 35%) is employed in management and science, but the difference between most and least employed is marginal for the remaining industries. Overall, the only increasing field over time is management and science whereas the other fields remained generally stagnant." 

#Creating visualization 3
merged_final_data <- merged_final_data |> 
  mutate(
    sector_type = fct_reorder(
      sector_type, 
      mean_percentage, 
      .desc = TRUE
      )
    ) |> 
  filter(!is.na(state_name))

#Creating visualization
merged_final_data |> 
  ggplot(aes(x = study_year, y = mean_percentage, fill = sector_type)) + 
  geom_density(stat = "identity", alpha = 0.5) +
  facet_wrap(~ state_name) +
  scale_x_continuous(breaks = c(2008, 2010, 2012, 2014, 2016, 2018)) +
  labs(
    x = "Year",
    y = "Population Percentage Employed",
    title = "Percentage of Population Employed in Sector by State",
    fill = "Sector"
    ) +
  theme_bw() +
  scale_fill_viridis_d(option = "B")
```

## Take-Away - Vis #1

-   Visualization #1 identifies states with highest scores in 4th-grade reading as MA, FL, NJ, and WY

    -   White students overrepresented in these states

    -   Schools must cater to cultural fluency of non-White students

## Take-Away - Vis #2

-   Visualization #2 illustrates NY, NJ, VT, and WY as states with highest government spending on students

    -   Warrants further research on public investment models in these states

## Take-Away - Regression Model

-   The best fit is in `regression_model_61`, with the predictors `family_poverty`, `ratio_of_one_to_both`, `average_center_child_care`, `average_family_child_care`, and `one_parent_households`

    -   Encourages support for m/paternity leave, subsidized childcare, and community support for families of low SES

## Take-Away - Vis #3

-   Visualization #3 suggests that states with high `reading_level` also have a high percentage of adults who work in "Management and Science"

    -   Prompts ideas for community-based vocational training for more to enter competitive fields

## Limitations

-   Only had the average 4th-grade `reading_level` per state, not county

-   Only had scores from 2022, so used community data from 2018, the most recent year of inputted data in `childcare_costs`

-   Standardized tests not perfect for aptitude, potential, or achievement measurement

    -   Consider disability or neurodiversity

## Works Cited

"EARLY SKILLS AND PREDICTORS OF ACADEMIC SUCCESS." Connecticut Government, Hanover Research, Nov. 2016, portal.ct.gov/-/media/SDE/ESSA-Evidence-Guides/Early_Skills_and_Predictors_of_Academic_Success #:\~:text.=Behavioral%20skills%20have.%20been %20found,behavioral%20skills%20are %20developing%20appropriately.Accessed 6 Dec. 2023.

"Education and Socioeconomic Status." American Psychological Association, 2017, www.apa.org/pi/ses/resources/publications/education. Accessed 6 Dec. 2023.

## Works Cited (cont'd)

"High-quality child care contributes to later success in science, math." American Psychological Association, 15 June 2023, www.apa.org/news/press/releases/2023/06/quality-child-care-science-math#:\~:text=Washington%20%E2%80%94%20Children%20 who%20receive%20high,by%20the%20American%20. Psychological%20Association. Accessed 6 Dec. 2023.

Perry-Jenkins, Maureen. "How a Parent's Experience at Work Impacts Their Kids." Harvard Business Review, 31 Jan. 2023, hbr.org/2023/01/how-a-parents-experience-at-work-impacts-their-kids#:\~:text=A%20multi%2Dyear%20study%20following,by%20 their%20parents%27% 20work%20lives.&text=It%27s%20no%20secret%20that%20our, our%20lives%20outside%20of%20work.

## Works Cited (cont'd)

"Racial and Ethnic Achievement Gaps." Stanford Center for Education Policy Analysis. Accessed 6 Dec. 2023.
