---
title: Proposal
author: Team Janíce
date: 11/6/2023
---

## Overview:

In this project, we seek to identify both community-level sociodemographic factors and state-level funding factors that predict statewide fourth-grade achievement rates in reading. Our ultimate goal is to determine pathways for policy-based intervention in the future to maximize educational success for youth of all backgrounds. We came upon this topic as a result of the content we are learning in our FOCUS cluster, Knowledge in the Service of Society. We are seeking to go beyond identifying community-level inequities and instead hope to investigate root causes and their logical solutions. 

To properly explore these factors, our project will include four major steps:

First, we will teach ourselves how to use the `maps` package to create a choropleth map of the United States where a gradient scale demonstrates the state achievement rates in fourth grade reading. This visualization will allow us, and others, to easily identify the states with highest achievement rates. From here, we can conduct further analysis on the funding distributions and community-level variables in these "best-practice states." 

Second, we will make a second choropleth map (a map that overlays a color gradient across a geogprahic region to portray the level of another variable across locations) using the `expenditures` dataset. This visualization will serve as a comparison to the first one, providing a visual representation of total school expenditures in elementary schools by state. By creating two maps of the same area (the United States) and using gradient scales to indicate both funding levels and achievement scores, viewers will be able to easily draw their own conclusions about possible associations between the two factors. In later parts of our project, we will use models to come to objective standpoints. 

Third, we will use forward selection (a systematic process of adding variables to a multiple linear regression model and comparing R-squared values) to identify community-level factors (e.g. poverty rate, median household income, average cost of childcare) that accurately predict statewide fourth grade achievement rates in reading. We will select potential predictors based on background knowledge from our other FOCUS classes and then move to statistical methods (adjusted-R2) to assess their importance in this context. Ultimately, we will have a linear regression model that merges a collection of community-level socioeconomic variables to predict fourth grade reading levels. After identifying which sociodemographic changes impact education achievement, we will be able to later theorize over possible steps to progress towards overall improvement via these specific pathways.

The fourth goal we want to accomplish is largely based upon the initial visualizations we create. Using the choropleth map, regression model, and R-squared analysis, we will ideally identify one or two noteworthy variables that appear to be the best predictors of achievement in fourth graders. We will complete 1-3 of these three options based on initial findings.

-   If we see that total elementary school expenditures seems to predict achievement rate based on the choropleth map from goal #2, we will analytically confirm this relationship through linear regression. We will then use the `mutate()` function to see whether the relationship is maintained based on the percentage of total state funding that is spent in elementary schools or whether an association exclusively exists between the total amount of elementary school funding and achievement rates.

-   Based on knowledge from our other KISS classes, we will explore the interactive effect between two significant community-level predictors of achievement rates to get a more granular understanding of how community structures combine with one another to predict academic success in youth.

-   We will determine a threshold for so-called "high achievement rates" and then subset our data to include states with achievement rates past this threshold. For these states, we will create bar charts that allow us to examine the distribution of funding in every sector within total state expenditures.

    -   We will use these bar charts to determine which sectors (e.g. transportation, Medicaid, etc.) seem to be getting the greatest allocation of state funding in top achievement states and then use multiple linear regression to confirm our hypothesized associations between funding distribution and achievement outcomes.

## Loading and Exploring Data

```{r}
#| label: loading-packages
#| message: false 

library(tidyverse)

```

```{r}
#| label: loading-datasets
#| message: false

#loading dataset childcare_costs
childcare_costs <- read_csv("data/childcare_costs.csv")

#loading dataset counties
counties <- read_csv("data/counties.csv")

#loading dataset expenditures
expenditures <- read_csv("data/state_expenditures.csv")

#loading dataset scores
scores <- read_csv("data/scores.csv")

```

The `expenditures` dataset is sourced from the Kaiser Family Foundation, through the National Association of State Budget Officers (NASBO)'s [State Expenditure Report Fiscal 2020-2022](https://www.nasbo.org/reports-data/state-expenditure-report). We chose this dataset because we wanted to compare government spending by both state and sector, focusing on 2021 educational expenditure amounts and their effects upon 2022 achievement rates from the `scores` dataset. The `expenditures` dataset has `r nrow(expenditures)` observations and `r ncol(expenditures)` variables. The observations are separated by `state` (with the additions of the District of Columbia and a row for the United States average) and specify 8 different types of public spending per state: `Elementary & Secondary Education`, `Medicaid`, `Transportation`, and `Higher Education`, `Public Assistance`, `Corrections`. 

The `scores` dataset is based on the Nation's Report Card data that uses the 2022 National Assessment of Educational Progress (NAEP) assessments with scores from 0-500 in mathematics, reading, writing, and sciences at grade 4. This dataset compares state and jurisdiction performance to the national public average scale score of 216. We chose this dataset because we need a quantitative way to analyze educational progress within each state/jurisdiction. The `scores` dataset has `r nrow(scores)` observations and `r ncol(scores)` variables. The observations are separated by state (plus the Department of Defense Education Activity, National Public, District of Columbia, and Puerto Rico). The dataset displays the average scores for each region (`MN`), the difference from the National Public (`SigDiff`), how significant the score is from the National Public (`SigSymbol`), the percentage at or above Basic (`AB`), and the percentage at or above Proficient (`AP`).

The `childcare_costs` dataset is from a federal source, the National Database of Childcare Prices. Measured from 2008 to 2018, its data is captured at the county level by for whom, what, where (, etc.) the childcare is provided. It also measures these data in conjunction with other seemingly unrelated variables and circumstances like teenage unemployment or young-parent labor. In the set, there are `r nrow(childcare_costs)` rows and `r ncol(childcare_costs)` columns. This dataset seemed pertinent because of how deeply intertwined it is with both the basis and offshooting considerations of our project. For example, it ties childcare costs with race, location, gender, income, and more; tying these considerations in with achievement rates provides a comprehensive look at predictors for childhood success. It is our hope that this dataset provides breadth to the analyses we proceed with conducting.

## Repository Organization:

Our repository is organized in a simple yet effective manner. All data files have been loaded into the data folder as csv files with a corresponding data dictionary in the README.md in the folder. Quarto files and the project itself have been left in the first level of the repository so that they have easy access to the data files and other objects.

## Weekly Timeline:

Week #3 (part 1---before 11/7) 

-   Finalize write-up of high-level goal #4 - Nush 

-   Create this "plan of attack" - Nush 

-   Finish exploring the datasets - Alan and Noah 

    -   Determine the \# of observations and variables (using inline code) in each dataset

    -   Identify and define the relevant variables in each dataset (using inline code) 

    -   Load all datasets into the project repository (3) 

-   Finalize the organization of our project repository - Julia 

    -   Create appropriate folders in the project repository with appropriate README.md files 

    -   Add/fix inline code in the proposal 

    -   Update the data dictionary 

-   Asynchronous discussion to assign each of the outlined responsibilities to various team members 

-   Upload a draft of the proposal into official project repository - Nush 

Week #3 (part 2---after 11/7) 

-   Team meeting to make adjustments to our proposal following peer review 

Week #4 

-   Merge the "expenditures" dataset with "scores" dataset - Noah 

-   Merge the "childcarecosts" dataset with "scores" dataset - Noah

-   Adapt all datasets to only include variables of interest - Noah 

-   Read through the description of  the "maps" package in R - Julia and Alan 

    -   Work collaboratively to create two choropleth maps through the functions in the "maps" package (visualizations #1 and #2) 

-   Use forward selection to build a multiple linear regression model as outlined in goal #3 and document the process (including adjusted R2 values) - Nush 

Week #5 

-   Team meeting: 

    -   Review feedback on proposal and make adjustments to our goals accordingly 

    -   Select which of the options from goal #4 to pursue 

    -   Create a mini-timeline for the rest of the week to asynchronously achieve our finalized goal #4  

-   Finish goal #4 as planned at team meeting 

Week #6: 

-   Interpret findings and create a draft of write-up

    -   Introduction - rationale for project and how it holistically connects to KISS - Noah

    -   Summary of goals and process - Noah

    -   Findings from visualization #1 (use inline code) - Alan 

    -   Findings from visualization #2 (use inline code) - Alan 

    -   Findings for visualization #3 (use inline code) - Alan

    -   Narrative for how we chose to approach goal #4 - Nush 

    -   Findings from visualization(s) #4 (use inline code) - Nush 

    -   Applications and recommendations for further research - Julia 

    -   Limitations - Julia/Noah (depending on how long the other steps take) 

Week #7: 

-   Team meeting #1: 

    -   Finalize write-up together 

    -   Use write-up as an outline to design presentation 

    -   Make adjustments to all parts of the project according to peer review

-   Team meeting #2: 

    -   Practice and refine presentation
